---
title: "Chapter 3"
output: html_notebook
---

## Describing Relationships
- In the last chapter we described the shape of, say, the temperature readings in the airquality dataset.
- In this chapter, we will be exploring whether there is a relationship between temperature and the month in which the temperature was taken (spoiler alert: there is!).

### Multivariate data
##### There are three broad classes of bivariate (or two variable) relationships: 
- The relationship between one categorical variable and one continuous variable 
- The relationship between two categorical variables 
- The relationship between two continuous variables

```{r}
head(iris)
```

```{r}
mean(iris$Petal.Length[iris$Species=="setosa"])
mean(iris$Petal.Length[iris$Species=="versicolor"])
mean(iris$Petal.Length[iris$Species=="virginica"])
```

There's a far easier way to do this:
```{r}
by(iris$Petal.Length, iris$Species, mean)
```

If we had reason to believe that the virginicas have a more widely varying petal length than the other two species, we could pass the sd function to the by function as follows:
```{r}
by(iris$Petal.Length, iris$Species, sd)
```

```{r}
summary(iris)
```

```{r}
by(iris$Petal.Length, iris$Species, summary)
```

Is a relationship between the average temperature across months?
```{r}
by(airquality$Temp, airquality$Month, mean)
```

```{r}
library(ggplot2)
ggplot(airquality, aes(x=as.factor(Month), y=Temp)) +
  geom_boxplot(fill='slateblue', alpha=0.2)
boxplot(Temp~Month, data=airquality, main="Relationship between Month and Temp")

```

### Relationship between two categorical variables

```{r}
ucba <- data.frame(UCBAdmissions)
head(ucba)
```


```{r}
cross <- xtabs(Freq ~ Gender+Admit, data=ucba)
cross
```

Is there a gender bias in UCB's graduate admissions process? Perhaps, but it's hard to tell from just looking at the 2x2 contingency table.

#### The values in each cell divided by the sum of the 4 cells:
```{r}
prop.table(cross)
```

#### The value of each cell divided by the sum of the row cells:
```{r}
prop.table(cross, 1)
```

#### The value of each cell divided by the sum of the column cells:

```{r}
prop.table(cross, 2)
```

To aid us in either implicating UCB of a sexist admissions machine or exonerating them, it would help to look at a proportions table. Using a proportions table, we can easily compare the proportion of the total number of males who were accepted versus the proportion of the total number of females who were accepted. If the proportions are more or less equal, we can conclude that gender does not constitute a factor in UCB's admissions process. If this is the case, gender and admission status is said to be conditionally independent.

```{r}
cross2 <- xtabs(Freq ~ Gender + Admit, data=ucba[ucba$Dept=="A",])
prop.table(cross2, 1)
```

Checking a different dept, "Dept A", there appears to be a bias towards women. If there were any bias in admissions, these data didn't prove it.


** This is probably the most famous case of Simpson's Paradox, and it is also why this dataset is built into R. The lesson here is to be careful when using pooled data, and look out for hidden variables. **

### The relationship between two continuous variables
```{r}
head(women)
```

This dataset has a positive relationship.

```{r}
library(ggplot2)
ggplot(women, aes(x=weight, y=height)) + 
  geom_point()
```

### Covariance

```{r}
cov(women$weight, women$height)
```

### Correlation coefficients

A Pearson correlation coefficient of 1 will denote a perfectly positive (linear) relationship, a r of -1 will denote a perfectly negative (linear) relationship, and a r of 0 will denote no (linear) relationship.
```{r}
cor(women$height, women$weight)
cor(women$height * 3.46, women$weight)
```
### Rank
When we use ranks instead, the pair that has the highest value on both the x and the y axis will be c(1,1), even if one variable is a non-linear function (cubed, squared, logarithmic, and so on) of the other. The correlations that we just tested will both have Spearman rhos of 1, because cubing a value will not change its rank.
```{r}
rank(c(8, 6, 7, 5, 3, 0, 9))
rank(c(8, 6, 7, 5, 3, -100, 99999))
```

```{r}
xs <- 1:100
cor(xs, xs+100)
cor(xs, xs^3)
```

```{r}
cor(xs, xs^3, method="spearman")
```

```{r}
iris.nospecies <- iris[, -5]
cor(iris.nospecies)
```

```{r}
cor(iris.nospecies, method='spearman')
```


### Categorical Continuous Variables


```{r}
library(ggplot2)
qplot(Species, Petal.Length, data=iris, geom="boxplot",fill=Species)
```


