---
title: "Homework 1"
author: Oleh Dubno
date: 01/25/18
output: html_notebook
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE,
                      message = FALSE)
```


*For all questions involving histograms, choose a sensible binwidth and breakpoints, unless otherwise indicated.*

### 1. Income


a) Describe in detail the features you observe in the box plots below, plotted with data from the *ex0525* data set, **Sleuth3** page. (see page 29 in *Graphical Data Analysis in R* for a list of features to concentrate on, and the numbered list on the bottom of page 43 for an example of how to describe features of a graph in words.) [5 points]

```{r}
# install.packages("Sleuth3")
library(Sleuth3)
library(tidyverse)

# convert Educ from an integer to a factor, and make "<12" the first factor level 
mydata <- ex0525 %>% 
  dplyr::mutate(Educ = forcats::fct_relevel(Educ, "<12"))

ggplot(mydata, aes(Educ, Income2005)) +
  geom_boxplot() + 
  coord_flip()   # for horizontal boxplots
```
* Income and the years of education have a positive relationship. As the years of education increase so does income. Variability in income increases as education goes up and the general mean is higher. Looking at 16>, there's a crazy outlier with $600,000 +. There's simply a trend of people earning more if they spend more time in school.

b) Plot a histogram of the `Income2005` variable in the data set referenced in part a). [3 points]
```{r}
ggplot(ex0525, aes(x=Income2005)) +
  geom_histogram(bins=60)
```

c) Use `+facet_wrap(~Educ)` to facet the histogram on education level. [3 points]
```{r}
ggplot(ex0525, aes(Income2005)) +
  geom_histogram(bins=60) +
  facet_wrap(~Educ)
```
d) What do you learn from the histograms that wasn't apparent in the box plots from question 1? [3 points]
* It's now way easier to see the distribution and where the data is concentrated. We could now see that there are more people who only completed 12 years of school. There's an inverse relationship between number of participants and their years of education. As the years of education increase the number of people participating goes down.

### 2. Respiratory Rates

a) Plot right closed and right open histograms for each of the two variables in the *ex0824* data set in the **Sleuth3** package using default binwidths and breaks. (4 histograms in total). [4 points]

```{r}
hist(ex0824$Age, left=TRUE, right=FALSE)
hist(ex0824$Age, left=FALSE, right=TRUE)
hist(ex0824$Rate, left=TRUE, right=FALSE)
hist(ex0824$Rate, left=FALSE, right=TRUE)
```

b) For which variable, `Age` or `Rate`, do the two versions differ more? Why? [3 points]
* The "Rate" histogram differs more, because the data is continuous as opposed to "Age" that's discrete. If right is "TRUE", which it is by default, the left is open and the right is closed eg (a, b] and vice versa if right is "FALSE", eg [a,b). 

* What this means for continuous data, if right="TRUE", is when a bin is (4, 8] and a continuous number 8.4 appears, it'll go into that bin. If a bin is [4,8) and a 8.4 appears, it'll go into the next bin. When graphing continuous data choosing which side you want to focus makes a difference.

| types      | limit           | open/closed  |
| ----------- |----------------| -----|
| (4, 8)      | 4 < x < 8      | left-open, right-open   |
| [4, 8)      | 4 ≤ x < 8      | left-closed, right-open |
| (4, 8]      | 4 < x ≤ 8      | left-open, right-closed |
| (4, 8]      | 4 ≤ x ≤ 8      | left-closed, right-closed |

c) Redraw the `Age` histograms with different parameters so that the right closed and right open versions are identical. [3 points]

```{r}
hist(ex0824$Age, breaks=2, left=TRUE, right=FALSE)
hist(ex0824$Age, breaks=2, left=FALSE, right=TRUE)
```

### 3. Movie budgets

Are there rounding patterns in the `budget` variable of the *movies* in the **ggplot2movies** package? If so, what are the patterns? (Note: according to the textbook this data set is in the **ggplot2** package, but it has since been moved to a separate package.) Support your conclusions with graphical evidence. You are encouraged to break the variable down into different budget ranges and consider them separately. [8 points] 

```{r}
# install.packages("ggplot2movies")
library(ggplot2movies)
```

```{r}
summary(m$budget)
```

```{r}
m$budget_group = cut(m$budget, c(0,100000,1000000, 10000000, 100000000, 1000000000))
```
```{r}
levels(m$budget_group) = c("1K","1M","10M", "100M", "1B")
```
```{r}
histogram(m$budget_group)
summary(m$budget_group)
```

* 53605 out of 58,788 movies have "N/A" for their budget. That leaves 5,183 with a recorded budget. Outside of the handful of movies that hit 1 billion in their budget, the rest of the movies don't appear to differ much in their budget. I would not be surprised if there are rounding errors.


### 4. Finches

a) Plot separate density histograms of the beak depth of the finches in *case0201* from the **Sleuth3** package, with density curves overlaid as on page 34 of the textbook. (However, do this by facetting on `Year` rather than using `grid.arrange`). [3 points]
```{r}
ggplot(case0201, aes(Depth)) +
  geom_density() +
  scale_fill_manual() +
  scale_x_continuous() +
  geom_histogram(alpha = 0.6, aes(y=..count../80/.1), binwidth=.1) +
  facet_grid(Year~.)
```

b) Plot both density curves on the same graph to facilitate comparison. Make 1976 yellow and 1978 blue.  Use alpha blending so the fills are transparent. [3 points]

```{r}
# converting Year into a factor
case0201$Year <- factor(case0201$Year, labels = c('1976', '1978'))
ggplot(case0201, aes(x=Depth, fill=Year)) +
  geom_density(alpha=.3) +
  scale_fill_manual(values = c("yellow","blue"))
```

c) Based on your graphs in parts a) and b), describe how the distributions differ by year. [3 points]
* We see a change from 1976 to 1978. Given this is a fair random sample for both years, the sample of birds from 1978 show less of the birds with smaller beak sizes. The density curve from 1978 has a higher concentration of birds with larger beak sizes. 

d) What is the cause of the difference according to the information in the help file? [3 points]

* The cause in the difference is drought. From 1976 to 1978 the common seeds that finches lived off of disappeared. The only seeds that survived the drought required finches to have bigger and stronger beaks; the ones that did not, couldn't reproduce.

### 5. Salary

Is the `Salary` variable in the *case0101* of **Sleuth3** normally distributed?  Use two different graphical methods to provide evidence. [6 points]
* case0101 {Sleuth3} is data on "Motivation and Creativity". There's no variable, `Salary`. Instead, I'll see if `Score`, a variable in the data set is normally distributed.

```{r}
summary(case0101$Score)
qqnorm(case0101$Score); qqline(case0101$Score, col = 2)
plot(density(case0101$Score))
```
* The `Score` variable is normally distributed. The mean and mode are very close together. The linear model shows that the data has a positive relationship. For the most part, the density plot shows evenness on both sides. 
