---
title: "Lecture 14"
output: html_notebook
---

```{r}
library(ggplot2)
```

```{r}
snow.url <- "https://www.ncdc.noaa.gov/snow-and-ice/daily-snow/NY-snowfall-201802.csv"
snow <- data.frame(read.csv(snow.url, skip=1))
```

```{r}
ggplot(snow, aes(x = Longitude, y = Latitude)) +
  geom_point() + theme_classic()
```

```{r}
snow <- snow %>% rename(Lat = Longitude, Lon = Latitude)
ggplot(snow, aes(x = Lon, y = Lat)) +
  geom_point() + theme_classic()
```

### Non visual summaries, such as: summary(), str(),  skimr::skim()
### When beginning multivariate analyses, work with a small number of variables first and a random sample of rows if the dataset is large
- start small, scatter plot; line plot (sales by year); gradually build up
### You do not need to use all the variables. Stay focused on the question(s) that interest you.
### p-hacking problem
- The “p-hacking” problem… is EDA immune?
- checking if a certain hypothesis, p is significant
- if the alpha value is .05 then that means 1/20 attempts at getting the p value, you'll get a false positive. Scientists will continue testing until they get one single significant test.
- only reporting significant level

```{r}
sales <- read.csv("https://community.watsonanalytics.com/wp-content/uploads/2015/08/WA_Sales_Products_2012-14.csv")
```

```{r}
# random 6 lines
index <- sort(sample(nrow(sales), 6))
sales[index,]
```

```{r}
countryorder <- sales %>% 
  filter(Year == 2012, `Product line` == "Personal") %>% 
  arrange(-TotalRev) %>% pull(`Retailer country`)
dflines <- snow %>% 
  mutate(`Retailer country` = fct_relevel(`Retailer country`, 
                                          as.character(countryorder)))
```

## Strategies for Spacial Data
- when to use maps and when not to use maps
- there's a cost to maps (we lose our axis)
- best utilized maps (John Snow, Cholera Map) - Geography is an independent variable
- location of where people live explains something
- different from graphing sport teams and put it on the map. Does location have to do with winning? Well it depends.
- is there a geographic patern to my data.
