---
title: "Lecture 8"
output: html_notebook
---

# Strategies
- Use techniques to deal with overplotting (alpha blending, open circles)
- Don't plot all points (remove outliers, subset data, sample data)
- Transform to log scale

## Look at Intervals separetely
```{r}
# install.packages("magrittr")
library(ggplot2)
library(ggplot2movies)
library(magrittr)
library(dplyr) 
```

```{r}
binnedmovies <- movies %>%
  mutate(mybin = ntile(votes, 10)) %>%
  dplyr::select(rating, votes, mybin)

head(binnedmovies)
```

# Log Scales
```{r}
ggplot(movies, aes(votes, rating)) +
  geom_point(alpha=.1) +
  scale_x_log10() +
  theme_gray(10)

```

# SoeedSki data(2011)
```{r}
# install.packages("GDAdata")
library(ggplot2)
library(GDAdata)
```

```{r}
g1 <- ggplot(SpeedSki, aes(Year, Speed))
g1 + geom_bin2d(binwidth=c(10, 15))
```
```{r}

```

```{r}
g1 + geom_hex()
```

```{r}
g1 + geom_density_2d(bins=5)
```

# Base graphics
```{r}
# install.packages("MASS")
# install.packages("SpeedSki")
library(MASS)
f1 <- kde2d(SpeedSki$Year, SpeedSki$Speed, n=100)
str(f1)
```

```{r}
image(f1)
```

```{r}
smoothScatter(SpeedSki$Year, SpeedSki$Speed)
```

# Calculate the kde, plot with ggplot2
- converts it to a very long data frame
```{r}
df <- con2tr(f1)
```
```{r}
ggplot(df, aes(x, y)) + 
  geom_contour()

ggplot(df, aes(x, y)) + 
  geom_tile(aes(Tile))
```

# Scatter plot matrices
```{r}
moveisdf <- data.frame(movies)
splomvar <- moveisdf %>% dplyr::select(length, rating, year, budget)
plot(splomvar)
```
```{r, fig.heaight=24, fig.width=16}
library(lattice)
splom(splomvar)
```
# Continuous + Categorical
## Mapping Options
- Continuous: x-axis, y-axis, color(no so great), size (not so great)
- Categorical: color, panels (rows, columns), shape (maybe)

```{r}
library(tidyverse)
ggplot(SpeedSki, aes)
```

```{r, fig.height=10, fig.width=8}
# install.packages("VGAMdata")
data(crime.us, package="VGAMdata")
crime.usR <- crime.us
names(crime.usR)[19:20] <- c("Larceny", "MotorVTheft")
ggpairs(crime.usR[, c(13:16, 18:20)],
        title="Crime rates in the USA",
        diag=list(continuous='density'), axisLabels='none')
```
```{r}

```

