---
title: "Lecture 5"
output: html_notebook
---

## Strip Charts
### see all the data.
### jitter (creates random noise)
- artificial noise, just to see the spread better.
## Strip Chart w/ boxplot
- great way to see summary statistics 
## Rug plot
- used on an axis
- an add on to a standard plot.
- gives the view of the two variables individually 
## Stem and leaf plot
- represents the first two numbers
## Dot Plot
- pretty much a histogram
## Normality
```{r}
x <- rnorm(1000, 50, 10)
qqplot(1:3, c(4, 4, 6))
```

```{r}
hist(x)
```
```{r}
df <- data.frame(x)
qqplot(df, aes(x)) + 
  geom_histogram(aes=(y= ..density..), fill='lightblue')
```
```{r}
qqnorm(x)
```
```{r}
x <- rexp(1000)
qqnorm(x)
qqline(x)
x <- rnorm(100)
qqnorm(x)
qqline(x, col='red')
```
## geom_histogram, geom_bar, geom_col
```{r}
library(tidyverse)
set.seed(5702)
df <- data.frame(x1=rnorm(100))
ggplot(df, aes(x1)) +
  geom_histogram()
```
```{r}
# has spaces between the bars
ggplot(df, aes(x1)) +
  geom_bar() +
  theme_gray(16)
```

```{r}
df2 <- data.frame(name=c("GA","QJ","AI"), count=c(31, 27, 27))
ggplot(df2, aes=(name, count))
```
```{r}
df <- data.frame(x=rnorm(100))
ggplot(df, aes(x)) +
  geom_histogram(fill="blue")
```
## Tidy Data
- Use package tidyr and gather to transform and clean up data

### Symptoms of messy data:
- no legend
- feel like you need grid.arrange()

```{r}
library(MASS)
head(painters)
```

```{r}
library(magrittr)
library(tidyverse)
tidypaint <- painters %>% rownames_to_column("Name") %>% 
  gather(key, value, -Name, -School)
```

```{r}
tidypaint <- painters %>% rownames_to_column("Name") %>% 
  gather(key=Skill, value=Score, -Name, -School)
```

```{r}
head(tidypaint)
```

### Use grid arrange to arrange charts

