---
title: "Lecture 10"
output: html_notebook
---

# Multivariate categorical data
- Bar Charts (Goto for categorical data)
- Cleveland dot plots
- Mosaic plots (parallen coordinate plots; difficult to understand)
- Fluctation diagrams
- Frequency data formats (coding)
- Drawing plots (coding)
- Occupational Mobility data

## Two continuous variables
1. for continuous data we make scatter plot

2. what about two categorical variables:
- don't use scatter plot
- stacked bar chart (not so good bc they don't line up)
- grouped bar chart (better bc everything lines up at the bottom)
- dependent variable is the fill color in the grouped bar chart
- grouped bar charts with facets Instead of color use variable names in chart)

3. cleveland dot plot(two variables)

# Three Categorical Variables
1. grouped bar chart with facet + color 
2. Cleveland dot plot
- grouped bar chart with facet and color 

# Chi Squared Test
http://www.ling.upenn.edu/~clight/chisquared.htm
- The Chi-square test is intended to test how likely it is that an observed distribution is due to chance.
- Likelihood that's proportional to the sizes of the group. The ratio affects the expected value.
- instead of using the frequency in data, it uses the ratio in the data.

- A Chi-square test is designed to analyze categorical data

# Mosaic Plot

# Mosaic plots (three variables)

# Mosaic Pairs Plot

# Doubledecker bar charts
- pretty much a mosaic plot

# Convert from one form to another
easy to get things into countr and tables
hard to go back to cases
cases -> counts -> tables
```{r}
Age
```

```{r}
data.frame(Titanic)
newct <- 
t <- chisq.test()
```
```{r}
cases %>% group_by(Age, Favorite) %>% summarize(Freq=n())
# counts
counts2 <- counts(cases) # ???
# dependent variable goes first; independent goes second
vcd::mosaic(Favorite~Age, counts2)
```

