---
title: "Lecture 11"
output: html_notebook
---


# Mosaic Plot
- Any filled rectangular plot (no white space) with consistent numbers of rows and columns, in which the area of each small rectangle is proportional to the frequency count for a unique combination of levels of the categorical variables displayed.

- github link: https://github.com/jtr13/codehelp/blob/master/R/mosaic.md

```{r}
library("grid"); library("vcd")
```

```{r}
csv_url <- url('https://raw.githubusercontent.com/jtr13/codehelp/master/data/MusicIcecream.csv')
df <- read.csv(csv_url)
```

Splitting on age
```{r}
mosaic(~Age, df)
```

Split on age then music
```{r}
mosaic(Music ~ Age, df)
```
Split on Age, then Music, then Favorite:

```{r}
mosaic(Favorite ~ Age + Music, df)
```
## Direction of splits

Note that in the previous example, the direction of the splits is as follows:

1. Age -- horizontal split

2. Music -- vertical split

3. Favorite -- horizontal split

This is the default direction pattern: alternating directions beginning with horizontal. Therefore we get the same plot with the following:

```{r}
mosaic(Favorite ~ Age + Music, direction=c("v","v","h"), df)
```
### Ff it's ordinal data, it should be ordered.
### The darked color should be at the botton.
```{r}
library(grid)
mosaic(Favorite ~ Age + Music, gp=gpar(fill=c("lightblue", "blue")), df)
```

```{r}
labeling <- labeling_border(rot_labels=c(45, -45, 0, 0))
mosaic(Favorite ~ Age + Music, labeling=labeling, df)
```

The rot_labels = vector sets the rotation in degrees on the four sides of the plot in this order: top, right, bottom, left. (Different from the typical base graphics order!) The default is rot_labels = c(0, 90, 0, 90).

### Abbreviate labels:
```{r}
mosaic(Favorite ~ Age + Music, 
       labeling = labeling_border(abbreviate_labs = c(3, 1, 6)), 
       df)
```
### Remove spacing
```{r}
mosaic(Favorite ~ Age + Music,
       spacing = spacing_equal(sp = unit(0, "lines")),
       df)
```

### Change border color

```{r}
mosaic(Favorite ~ Age + Music,
       gp = gpar(fill = c("lightblue", "blue"),
                 col = "white"),
       spacing = spacing_equal(sp = unit(0, "lines")),
       df)
```

# Tree map
- filled rectangular plot representing hierarchical data (fill color does not necessary represent frequency count)
- these are nests and does not tell you anything about the relationship between the variables

# Spine plot
- Have lines that go all the way across without horizontal splits

# Bar Plot vs Mosaic Plot
relative frequency bar plot is pretty much the same as a mosaic plot with fixed bin length.
"i feel that is not a mosaic plot unless the bins are uneven"

# Likert Data ( Ordinal Data *not discrete* )
### E.g.
- strongly agree
- agree
- don't know
- disagree
- strongly disagree


## Diverging Stacked bar Chart 

```{r}
# install.packages("HH")
?? HH::likert
```

# Heatmaps

better going from nominal to categorical

```{r}
library(gridExtra)
library(ggplot2)
x <- 1:3
y <- c(5, 2, 7)
df <- data.frame(x, y)
g1 <- ggplot(df, aes(x, y)) + geom_point()
g2 <- ggplot(df, aes(x, y)) + geom_tile()
grid.arrange(g1, g2, nrow=1)
```
```{r}
df$w <- c(1.4, .4, .2)
df$h <- c(.5, 1.3, .8)

ggplot(df, aes(x, y)) + geom_tile(aes(width=w, height=h))
```
# Use factor when working with categories to avoid ordinal data that are intergers from becoming gradients.
