---
title: "Lecture 13"
output: html_notebook
---

# Missing Data
```{r}
library(skimr)
library(ggplot2)
library(magrittr)
library(mi)
```

### Looking for patterns in missing data (row/ column patterns)
- purely obeservation descriptive
- do all variables have the same percentage of values missing?
- Are there row patterns to the missing values?
- use a heat map to visualize 
- with heat maps you'll better see missing data
- heatmap - geom_tile()

```{r}
mycars <- mtcars
skimr::skim(mycars)%>%filter(stat=="missing")%>%arrange(desc(value))%>%select(variable,type,value)%>%mutate(percent=value/nrow(mycars))
```

### Skimr gives a good summary of any data frame
- great summaries
- 
```{r}
skimr::skim(mtcars)
```
### MI (Missing data frame function)

```{r}
mi::missing_data.frame(mtcars)
```
```{r}
x <- mi::missing_data.frame(mycars)
show(x)
heatmap(x)
```
## Tow/column missing patters
- Patters 

```{r}
x@patterns

```

# Visna 
- shows patterns not the rows
- what do the patterns mean?
- less rows and each row represents a missing pattern
- Also has a bar to show how much data is missing
- visna()

# Loking for patterns in missing data
- value patterns
- are missing patterns correlated with values of another variable? (MAR=missing at random)
- are certain (true) value ranges more likely to be missing? (MNAR = missing not at random)
- there's a reason certain data is missing. research and figure out why certain values are missing

```{r}
df <- read.csv("https://raw.githubusercontent.com/jtr13/dataquality/master/SAT2010.csv")
df
```

```{r}
x <- mi::missing_data.frame(df)
x
```

```{r}
head(df)
```

## MAR (Missing At Random)
## MNAR (Missing Not At Random)

# Does the missing data vary by a certain column (borough?)
- Use the dbn column to identify the borough.
```{r}
df$DBN2 <- gsub('[0-9]','',df$DBN)
```
```{r}
df
```


https://www.rstudio.com/resources/webinars/extracting-data-from-the-web-part-1/
rvest for scrapping? 